---
title: "nba draft helper"
author: "Nathan Hwangbo"
date: "12/18/2020"
output: html_document
---

small script to get familiar with r nba package

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


load libraries
```{r libs, warning = F, message = F}
#devtools::install_github('abresler/nbastatR')
library(nbastatR)
library(tidyverse)
```


```{r dat}

#2017-2020 season. training dat
players20172020 <- bref_players_stats(seasons = 2018:2020, 'per_game') %>%
  select(namePlayer, groupPosition, slugPosition, yearSeason, agePlayer, 
         slugTeamBREF, countGamesStarted, ends_with("PerGame"))

# get new rosters
rosters2021 <- seasons_rosters(2021) %>%
  select(namePlayer, slugTeam)

train_df <- players20172020 %>%
  left_join(rosters2021, by = 'namePlayer') %>%
  mutate(team = ifelse(is.na(slugTeam), slugTeamBREF, slugTeam)) %>%
  select(-slugTeam, -slugTeamBREF) %>%
  mutate(yahoo = ptsPerGame + 1.2*trbPerGame + 1.5*astPerGame + 
           3*stlPerGame + 3*blkPerGame - tovPerGame) %>%
  arrange(desc(yahoo)) %>%
  # player summary
  group_by(namePlayer) %>%
  mutate(is_improve = yearSeason[1] == 2019) %>%
  ungroup() %>%
  # team summary
  group_by(team, yearSeason) %>%
  mutate(team_med_yahoo = median(yahoo)) %>%
  ungroup()
  
  

  


```


Team quality
```{r}
train_df %>% 
  distinct(team, yearSeason, team_med_yahoo) %>% 
  arrange(yearSeason, desc(team_med_yahoo)) %>% 
  ggplot() + 
  geom_col(aes(team, team_med_yahoo)) + 
  facet_wrap(~yearSeason) +
  coord_flip()
```


identify teams with 1 star

```{r}
train_df %>%
  group_by(team, yearSeason) %>%
  summarize(first_diff = yahoo[1] - yahoo[2]) %>%
  arrange(desc(first_diff))
  

```



category scoring setup
```{r}
# only take categories of interest

cat_df <- train_df %>%
  filter(minutesPerGame > 20) %>%
  transmute(name = namePlayer, pos = groupPosition, year = yearSeason, agePlayer, countGamesStarted, 
            num_min = minutesPerGame,
            ptsPerGame, trbPerGame, astPerGame, stlPerGame, blkPerGame, fg3mPerGame,
            ftpPerGame = ifelse(ftaPerGame > 0, ftmPerGame / ftaPerGame, 0),
            fgpPerGame = ifelse(fgaPerGame > 0, fgmPerGame / fgaPerGame, 0),
            tovPerGame
            ) %>%
  pivot_longer(ends_with('PerGame'), 
               names_to = 'stat',
               values_to = 'pergame')


```

best 5 players by category
```{r}
cat_df %>%
  group_by(stat, year) %>%
  slice_max(pergame, n = 25) %>%
  ungroup() %>%
  group_by(name) %>%
  mutate(is_tov = ifelse(any(stat == 'tovPerGame'), 1, 0),
         num_years = length(unique(year))) %>%
  group_by(name, year) %>%
  mutate(num_cat = n() - max(is_tov)) %>%
  ungroup() %>%
  arrange(desc(num_cat), name, desc(year)) %>%
  filter(num_cat >=3) %>%
  View

```


category breakdown -- how likely am i to win each of the columns?
```{r}
cat_df %>%
  group_by(year, stat) %>%
  mutate(statscale = (pergame - min(pergame)) / (max(pergame) - min(pergame))) %>%
  summarize(stat_var = var(statscale)) %>%
  ggplot() + 
  geom_col(aes(stat, stat_var)) +
  coord_flip() + 
  facet_wrap(~year)

```

```{r}
cat_df %>%
  group_by(year, stat) %>%
  mutate(pergamescale = (pergame - min(pergame)) / (max(pergame) - min(pergame))) %>%
  ungroup() %>%
  ggplot() + 
  geom_histogram(aes(pergamescale)) + 
  facet_grid(year~stat)
```




Question 1. which category is most consistent? e.g. if the ranking of steal leaders is consistent every year, that means that picking a high ranking stealer is beneficial


Question 2. how big are the differences in ranking? e.g. if one player leads steals every year consistently, it might not matter if the difference is only ~1 steal per game off. 

So we want to emphasize players who dominate categories consistently, where "domination" must be meaningfully large for fantasy points.







Idea: sample 100,000 fantasy teams by randomly selecting players (say, 2018-19). For each team, record the score for each of the categories. See which players consistently appear on winning teams

```{r}
# given a year, create a sample of 10 players and see how well they do
# input: year
# output: summed performance in each elague.
sample_team <- function(year){
  roster <- train_df %>% 
    filter(yearSeason == year, minutesPerGame > 20) %>%
    sample_n(10, replace = F)
  
  
  roster_long <- roster %>%
    transmute(name = namePlayer, pos = groupPosition, year = yearSeason, agePlayer, countGamesStarted, 
            num_min = minutesPerGame,
            ptsPerGame, trbPerGame, astPerGame, stlPerGame, blkPerGame, fg3mPerGame,
            ftpPerGame = ifelse(ftaPerGame > 0, ftmPerGame / ftaPerGame, 0),
            fgpPerGame = ifelse(fgaPerGame > 0, fgmPerGame / fgaPerGame, 0),
            tovPerGame
            ) %>%
  pivot_longer(ends_with('PerGame'), 
               names_to = 'stat',
               values_to = 'pergame') %>%
    arrange(stat)
  
  # put it in a single vector so we can stretch it out later lol
  agg_stats <- roster_long %>%
    group_by(stat) %>%
    summarize(total = sum(pergame), .groups = 'drop') %>%
    ungroup() %>%
    pull(total) %>%
    round(2) %>%
    paste(collapse = "__")
  
  roster_long %>%
    mutate(stats = agg_stats)
}

team_sim <- tibble(runs = 1:100000,
                   team = map(runs, ~sample_team(2019)),
                   players = map(team, ~unique(.x$name)),
                   stats = map_chr(team, ~unique(.x$stats))
                   ) %>%
  separate(stats, into = sort(unique(cat_df$stat)), sep = "__", remove = T, convert = T) %>%
  mutate(across(ends_with('PerGame'), ~rank(-.x))) %>%
  mutate(sum_rank = reduce(select(., ends_with('PerGame')), `+`))




```

get the best players on avg (2019-2020 season)
```{r}
best2019 <- team_sim %>%
  unnest(players) %>%
  group_by(players) %>%
  summarize(n_sampled = length(unique(runs)),
            med_sumrank = median(sum_rank),
            .groups = 'drop'
            ) %>%
  arrange(med_sumrank)

```


how does this compare to 2020? are the best players consistent?

```{r}
team_sim2020 <- tibble(runs = 1:100000,
                   team = map(runs, ~sample_team(2020)),
                   players = map(team, ~unique(.x$name)),
                   stats = map_chr(team, ~unique(.x$stats))
                   ) %>%
  separate(stats, into = sort(unique(cat_df$stat)), sep = "__", remove = T, convert = T) %>%
  mutate(across(ends_with('PerGame'), ~rank(-.x))) %>%
  mutate(sum_rank = reduce(select(., ends_with('PerGame')), `+`))

```

```{r}
best2020 <- team_sim2020 %>%
  unnest(players) %>%
  group_by(players) %>%
  summarize(n_sampled = length(unique(runs)),
            med_sumrank = median(sum_rank),
            .groups = 'drop'
            ) %>%
  arrange(med_sumrank)

```


compile 2019, 2020 data to see if they're consistent
```{r}
best1920 <- best2020 %>%
  rename(n_sampled2020 = n_sampled, med_sumrank2020 = med_sumrank) %>%
  inner_join(best2019, by = 'players') %>%
  mutate(across(starts_with('med'), ~ rank(.x)))


best1920 %>%
  summarize(rank_cor = cor(med_sumrank, med_sumrank2020))  
```

now that we know that results are pretty consistent, let's use a weighted average of the two ranks to complete the analysis.
```{r}
# first, get a list of player positions
player_pos <- train_df %>%
  separate_rows(slugPosition, sep = '-') %>%
  distinct(namePlayer, slugPosition) %>%
  group_by(namePlayer) %>%
  mutate(num_pos = n()) %>%
  ungroup() %>%
  rename(players = namePlayer, pos = slugPosition)

draft_selector <- best1920 %>%
  mutate(avg_rank = 0.55 * med_sumrank2020 + 0.45*med_sumrank) %>%
  arrange(avg_rank) %>%
  select(players, avg_rank) %>%
  left_join(player_pos, by = 'players')
```

draft idea: to start the draft, look at the overall list. to start filling out positions, look at the position specific dataframes
```{r}

draft_selector

draft_selector %>%
  group_by(pos) %>%
  slice_min(avg_rank, n = 20) %>% 
  group_split()
```
